Model Search in Relational Domains

### 2016-10-21 algorithm.tex
sync

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKw{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{conflict}
\SetKwData{safe}{safe}
\SetKwData{sat}{sat}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{%An abstract domain $A$, 
A program $\mathcal{S}$, a
  propagation heuristics $H$, a decision heuristics $\mathit{decide}$,
and a conflict analysis heuristics $\mathit{conflict\_analysis}$.\rmcmt{calling
them heuristics is not a good idea since these are key phases of the cdcl
algorithm. Each phase may have a heuristics for some sub phase but the algorithm
of each phase still remians a procedure and not a heuristic}}
\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
$\mathcal{T} \leftarrow \langle\rangle$ \;
$\mathcal{R} \leftarrow \emptyset$ \;
$\mathit{result} \leftarrow \mathit{deduce}(S,\mathcal{T},\mathcal{R}, H)$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return \unsafe}
  $d \leftarrow \mathit{decide}(\mathit{abs}(\mathcal{T}))$ \;
%  $v \leftarrow v \meet d$ \; 
  $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{decision} \;
  $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$\;
  \Do{$\mathit{result} = \conflict$} {
%    learn() \;
    \lIf{$\neg \mathit{conflict\_analysis}(\mathcal{S},\mathcal{T},\mathcal{R})$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $ACDCL(A)$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The algorithm takes a program \pscmt{make this a set of transformers
  (parametrised by the propagation heuristics)}
$\mathcal{S}$ and propagation, decision, and conflict analysis
heuristics as inputs.  As internal state, it maintains a propagation
trail $\mathcal{T}$ and a reason trail \pscmt{map?} $\mathcal{R}$.
The propagation trail is an initially empty sequence of meet
irreducibles inferred by the abstract model search phase (deductions
and decisions). The reason trail maps \pscmt{certain?} elements of the
propagation trail to transformers $s\in\mathcal{S}$ that were used to
derive them. The \emph{abstract value} $\mathit{abs}(\mathcal{T})$
corresponding to the propagation trail $\mathcal{T}$ is the
conjunction of the meet irreducibles on the trail:
$\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with
$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty
sequence $\langle\rangle$.

@@

\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\mathcal{T}$?} The state of a ACDCL solver is a tuple 
of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
is a sequence of labelled information of the form $(m,s)$ where 
$m$ is a meet irreducible and $s = \mathsf{decision}$ if $m$ is a decision, 
or $s = \mathsf{deduction}$ if $m$ is inferred by a deduction. And $S$ is 
a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 
stores variable assignments of the form $(p, t)$, where $p$ is 
a propositional variable that appears at most once in the trail 
and $t$ is a truth assignment (true or false).  Whereas, a trail 
$\mathcal{T}$ in ACDCL contains a sequence of meet irreducibles 
inferred by deduction or decision phase where a variable in trail 
can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
bounds \pscmt{that's specific for a particular domain}.  
\rmcmt{define trail refinement}
\Omit {
A current valuation $v$ is a meet of all elements of trail
$\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
$m_i \in \mathcal{T}$.  A solver is in conflict if some clauses in
$\mathcal{S}$ is not abstractly satisfied by $v$.
}
%\pscmt{define ``consistent with trail''}

>>>>

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKw{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{conflict}
\SetKwData{safe}{safe}
\SetKwData{sat}{sat}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{%An abstract domain $A$, 
A program $\mathcal{S}$, a
  propagation heuristics $H$, a decision heuristics $\mathit{decide}$,
and a conflict analysis heuristics $\mathit{conflict\_analysis}$.\rmcmt{calling
them heuristics is not a good idea since these are key phases of the cdcl
algorithm. Each phase may have a heuristics for some sub phase but the algorithm
of each phase still remians a procedure and not a heuristic}}
\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
$\mathcal{T} \leftarrow \langle\rangle$ \;
$\mathcal{R} \leftarrow \emptyset$ \;
$\mathit{result} \leftarrow \mathit{deduce}(S,\mathcal{T},\mathcal{R}, H)$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return \unsafe}
  $d \leftarrow \mathit{decide}(\mathit{abs}(\mathcal{T}))$ \;
%  $v \leftarrow v \meet d$ \; 
  $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
  $\mathcal{R}[|\mathcal{T}|] \leftarrow \top$ \;
  $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$\;
  \Do{$\mathit{result} = \conflict$} {
%    learn() \;
    \lIf{$\neg \mathit{conflict\_analysis}(\mathcal{S},\mathcal{T},\mathcal{R})$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $ACDCL(A)$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The algorithm takes a program \pscmt{make this a set of transformers
  (parametrised by the propagation heuristics)}
$\mathcal{S}$ and propagation, decision, and conflict analysis
heuristics as inputs.  As internal state, it maintains a propagation
trail $\mathcal{T}$ and a reason trail \pscmt{map?} $\mathcal{R}$.
The propagation trail is an initially empty sequence of meet
irreducibles inferred by the abstract model search phase (deductions
and decisions). The reason trail maps \pscmt{certain?} elements of the
propagation trail to transformers $s\in\mathcal{S}$ that were used to
derive them. The \emph{abstract value} $\mathit{abs}(\mathcal{T})$
corresponding to the propagation trail $\mathcal{T}$ is the
conjunction of the meet irreducibles on the trail:
$\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with
$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty
sequence $\langle\rangle$.
\pscmt{trail refinement}

@@

%\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\mathcal{T}$?} The state of a ACDCL solver is a tuple 
%of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
%is a sequence of labelled information of the form $(m,s)$ where 
%$m$ is a meet irreducible and $s = \mathsf{decision}$ if $m$ is a decision, 
%or $s = \mathsf{deduction}$ if $m$ is inferred by a deduction. And $S$ is 
%a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 
%stores variable assignments of the form $(p, t)$, where $p$ is 
%a propositional variable that appears at most once in the trail 
%and $t$ is a truth assignment (true or false).  Whereas, a trail 
%$\mathcal{T}$ in ACDCL contains a sequence of meet irreducibles 
%inferred by deduction or decision phase where a variable in trail 
%can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
%bounds \pscmt{that's specific for a particular domain}.  
%\rmcmt{define trail refinement}
%\Omit {
%A current valuation $v$ is a meet of all elements of trail
%$\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
%empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
%$m_i \in \mathcal{T}$.  A solver is in conflict if some clauses in
%$\mathcal{S}$ is not abstractly satisfied by $v$.
%}
%\pscmt{define ``consistent with trail''}

### 2016-10-21 domains.tex
sync

For notational convenience we will denote elements of
$\mathcal{B}^b$ by their concretisations to propositional formulae,
e.g.\ for the boolean variable vector $(g_1,c)$ \rmcmt{what is g1, what is c?} 
with abstract value $(\false,\top)$ we will write $\neg g_1$.

%-------------------------------------------------------------------------------
\paragraph{Template Polyhedra Abstract Domain}
%
The template polyhedra domain~\cite{sriram} represents sets $X$ of
valuations of a vector of variables $\vec{x}$ \rmcmt{not clear what is a
valuation?} such that
$\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
and a constant vector $\vec{d}$. The domain of $\vec{d}$ is augmented
by a special element to denote the minimal element $\bot$ of the lattice.
%
The abstraction function is defined as $\alpha(X) = \min \{\vec{d}\mid
\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is
applied component-wise. There are several optimisation-based
techniques for computing the domain
operations~\cite{sriram,GS07b,BJKS15}\rmcmt{like what operations?}. In our implementation we use
the strategy iteration approach available in 2LS~\cite{BJKS15}.
%
The concretisation $\gamma(\vec{d})$ is the set $\{\vec{x}\mid
\mat{C}\vec{x}\leq\vec{d}\}$; $\gamma(\bot)=\emptyset$.
%
For notational convenience we will use conjunctions of linear
inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
of the \rmcmt{template polyhedra ?} domain. $\true$ is $\top$ and $\false$ is $\bot$.

@@

The interval domain is a non-relational domain because a single
inequality only contains a single variable.
%
The octagon domain, however, is relational.  A fundamental difference
between relational and non-relational domains is that relational
domains require the computation of a \emph{closure} in order to obtain
a normal form that is usually required for efficient \rmcmt{or precise?} 
domain operations.  An example of a closure computation for octagonal 
inequalities is shown below.
\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
The closure computes all implied domain constraints.  For octagons,
the closure operation is very expensive (cubic complexity in the
number of program variables).
%
We will see that our model search algorithm performs closure computations lazily,
postponing them to the point where they are actually necessary.\rmcmt{are we
showing this ?}

@@

%-------------------------------------------------------------------------------
\paragraph{Abstract Transformers}
%
%In a typical abstract interpretation based Galois-connection setting
%over an over-approximate domain, every concrete element has a unique
%over-approximate representation in the abstract.  Likewise, every
%concrete transformer is over-approximated by a unique abstract
%transformer.  We now define an abstract deduction transformer.
%
An abstract transformer transforms an abstract value $a$ through a
statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
The best abstract transformer is \rmcmt{is this equation updated ?}
\[\llbracket s
\rrbracket^\sharp(a)=\bigsqcup \{a'\mid a'\sqsubseteq a, a'\models s\}\] where we write
$a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
of $a$. Any abstract transformer that over-approximates the least
upper bound is a sound transformer and can be used in our algorithm.

>>>>

\pscmt{In Figure~\ref{}}
For notational convenience we will denote elements of
$\mathcal{B}^b$ by their concretisations to propositional formulae,
e.g.\ for the boolean variable vector $(g_1,c)$ \rmcmt{what is g1, what is c?} 
with abstract value $(\false,\top)$ we will write $\neg g_1$.

%-------------------------------------------------------------------------------
\paragraph{Template Polyhedra Abstract Domain}
%
The template polyhedra domain~\cite{sriram} represents sets $X$ \pscmt{split this up into two sentences} of
valuations of a vector of variables $\vec{x}$ \rmcmt{not clear what is a
valuation?} such that
$\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
and a constant vector $\vec{d}$. The domain of $\vec{d}$ is augmented
by a special element to denote the minimal element $\bot$ of the lattice.
%
The abstraction function is defined as $\alpha(X) = \min \{\vec{d}\mid
\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is
applied component-wise. There are several optimisation-based
techniques for computing the domain
operations, such as $\sqcap$, $\sqcup$, etc~\cite{sriram,GS07b,BJKS15}\rmcmt{like what operations?}. In our implementation we use
the strategy iteration approach available in 2LS~\cite{BJKS15}.
%
The concretisation $\gamma(\vec{d})$ is the set $\{\vec{x}\mid
\mat{C}\vec{x}\leq\vec{d}\}$; $\gamma(\bot)=\emptyset$.
%
For notational convenience we will use conjunctions of linear
inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
of the template polyhedra domain. $\true$ is $\top$ and $\false$ is $\bot$.
\pscmt{give $\mat{C}$ and $\vec{d}$ for this example}

@@

The interval domain is a non-relational domain because a single
inequality only contains a single variable.
%
The octagon domain, however, is relational.  A fundamental difference
between relational and non-relational domains is that relational
domains require the computation of a \emph{closure} in order to obtain
a normal form that is usually required for precise and efficient 
domain operations.  An example of a closure computation for octagonal 
inequalities is shown below.
\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
The closure computes all implied domain constraints.  For octagons,
the closure operation is very expensive (cubic complexity in the
number of program variables).
%
We will see that our model search algorithm performs closure computations lazily,
postponing them to the point where they are actually necessary.\rmcmt{are we
showing this ?}

@@

%-------------------------------------------------------------------------------
\paragraph{Abstract Transformers}
%
%In a typical abstract interpretation based Galois-connection setting
%over an over-approximate domain, every concrete element has a unique
%over-approximate representation in the abstract.  Likewise, every
%concrete transformer is over-approximated by a unique abstract
%transformer.  We now define an abstract deduction transformer.
%
An abstract transformer transforms an abstract value $a$ through a
statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
The best abstract transformer is \rmcmt{is this equation updated ?}
\[\llbracket s
\rrbracket^\sharp(a)=\bigsqcap \{a'\mid a'\sqsubseteq a, a'\models s\}\] where we write \pscmt{fix this}
$a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
of $a$. Any abstract transformer that over-approximates the least
upper bound is a sound transformer and can be used in our algorithm.
